---
title: ""
author: ""
date: ""
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE)
```

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}

code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r}
library(dataQualityR)
library(e1071)
library(mice)
library(corrplot)
library(ggplot2)
library(ggpubr)
library(scales)
library(caret)
library(dplyr)
library(tidyverse)
library(sf)
library(gganimate)
library(MASS)
library(VIM)
library(glmnet)
```

# Outline:


# 1. Introduction

# 2. Data Preparation

## 2.1. Data collection

## 2.2. Data cleaning

- Data reading
```{r}
df<-read.csv("Data1.csv")
dim(df)
```

- Data overview
```{r}
table(df$Arrests)
```
- Remove School.ID column
```{r}
df$School.ID<-NULL # Order column looks like meaningless
```

- having a data report for cat and num 
```{r}
num.file <- paste(getwd(), "/dqames_num.csv", sep= "")
cat.file <- paste(getwd(), "/dqames_cat.csv", sep= "")
checkDataQuality(data= df, out.file.num= num.file, out.file.cat= cat.file)
```

- Checking missing values
```{r}
library(VIM)
aggr_plot <- aggr(df, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(df), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
```

```{r}
dim(df)
```

- remove columns with missing values
```{r}
df$LawEnforcementSchool <- NULL
df$LawEnforcementActivities <- NULL
df$LawEnforcementProblemParticipation<- NULL
df$LawEnforcementPreventionParticipation <- NULL
df$LawEnforcementStudentMentoring<- NULL
df$LawEnforcementTeaching<- NULL
df$LawEnforcementReporting<- NULL
df$LawEnforcementDisciplinePolicies<- NULL
df$LawEnforcementArrestPolicies<- NULL
df$LawEnforcementReportingPolicies<- NULL
```

```{r}
table(df$Arrests)
```

- Group school's policies
```{r}
df$SchoolSecurity <- rowSums(df[,c(1:8, 14, 16:20,22)])
df$SchoolProceduresRules <- rowSums(df[,c(9:13, 15, 21, 23:25)])
df$SchoolPreventionProgram <- rowSums(df[,c(25:41)])
df$SchoolMentalHealthServices <- rowSums(df[,c(42:43)])
df$SchoolTrainingProgram <- rowSums(df[,c(44:53)])
```

- remove a group of dummies
```{r}
df[,2:53] <- NULL
dim(df)
str(df)
summary(df)
```

- Write out new data
```{r}
write.csv(df,"data2.csv", row.names = FALSE)
```