---
title: "Project_Team_10_DataCleaningPart2 + Modeling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE)
```

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}

code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>


```{r}
library(dataQualityR) #for data cleaning
library(e1071) #SVM Training & Testing Models
library(mice) #deal with missing data
library(corrplot) #plot for correlation matrix
library(ggplot2) #visualization and plots
library(ggpubr) #customizing ggplot2
library(scales) #graphical scales map data to aesthetics
library(caret) #deal with Classification And REgression Training CART
library(dplyr) #data manipulation: filter and arrange
library(tidyverse) #data manipulation
library(sf) #plot mapping
library(gganimate) #static visualization
library(MASS) #deal with data set
library(VIM) #tools for the visualization of missing or imputed values
library(glmnet) #Lasso and Elastic-Net Regularization
```

## Additional Cleaning for data2
- Read new data
```{r}
df2 <- read.csv("data2.csv")
```

- having a data report for cat and num 
```{r}
num.file <- paste(getwd(), "/dqames_num2.csv", sep= "")
cat.file <- paste(getwd(), "/dqames_cat2.csv", sep= "")
checkDataQuality(data= df2, out.file.num= num.file, out.file.cat= cat.file)
```

- change subclass from num to cat 
```{r}
df2[,5:13] <- lapply(df2[,5:13], as.factor)
df2[,20:21] <- lapply(df2[,20:21], as.factor)
df2[,40:43] <- lapply(df2[,40:43], as.factor)
```

```{r}
df2[,23:24]<-NULL # Order column looks like meaningless
df2[,24:35]<-NULL # Order column looks like meaningless
```

```{r}
str(df2)
```

# 3. Exploratory Data Analysis - EDA

```{r}
h1 <- hist(df2$TotalIncidents)
h2 <- hist(log(df2$TotalIncidents))

q1 <- ggplot(df2, aes(sample=TotalIncidents)) +
  stat_qq(color="dodgerblue4") + 
  stat_qq_line(color="red") +
  scale_y_continuous(labels=function(y){y/10^6}) +
  labs(title="QQ Plot for TotalIncidents", y="Ordered Values") +
  theme(plot.title=element_text(hjust=0.5))
q1

q2 <- ggplot(df2, aes(sample=log(TotalIncidents))) +
  stat_qq(color="dodgerblue4") + 
  stat_qq_line(color="red") +
  scale_y_continuous(labels=function(y){y/10^6}) +
  labs(title="QQ Plot for Log TotalIncidents", y="Ordered Values") +
  theme(plot.title=element_text(hjust=0.5))
q2
```

## 3.1. 1D
```{r}
h1 <- ggplot(df2, aes(SchoolSecurity))+ 
  geom_histogram() + 
  labs(title = "Histogram of SchoolSecurity") + 
  theme(plot.title=element_text(hjust=0.5)) +
  geom_vline(aes(xintercept=mean(SchoolSecurity)), color="blue", linetype="dashed", size=1)
h1

h2 <- ggplot(df2, aes(SchoolProceduresRules))+ 
  geom_histogram() + 
  labs(title = "Histogram of SchoolProceduresRules") + 
  theme(plot.title=element_text(hjust=0.5)) +
  geom_vline(aes(xintercept=mean(SchoolProceduresRules)), color="blue", linetype="dashed", size=1)
h2

h3 <- ggplot(df2, aes(SchoolPreventionProgram))+ 
  geom_histogram() + 
  labs(title = "Histogram of SchoolPreventionProgram") + 
  theme(plot.title=element_text(hjust=0.5)) +
  geom_vline(aes(xintercept=mean(SchoolPreventionProgram)), color="blue", linetype="dashed", size=1)
h3

h4 <- ggplot(df2, aes(SchoolTrainingProgram))+ 
  geom_histogram() + 
  labs(title = "SchoolTrainingProgram") + 
  theme(plot.title=element_text(hjust=0.5)) +
  geom_vline(aes(xintercept=mean(SchoolTrainingProgram)), color="blue", linetype="dashed", size=1)
h4

h5 <- ggplot(df2, aes(SchoolMentalHealthServices))+ 
  geom_histogram() + 
  labs(title = "Histogram of SchoolMentalHealthServices") + 
  theme(plot.title=element_text(hjust=0.5)) +
  geom_vline(aes(xintercept=mean(SchoolMentalHealthServices)), color="blue", linetype="dashed", size=1)
h5
```

## 3.2. 2D

```{r}
s1 <- ggplot(df2, aes(x=SchoolSecurity, y=log(TotalIncidents+1))) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="SchoolSecurity vs. y=log(TotalIncidents+1)") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))
s1

s2 <- ggplot(df2, aes(x=SchoolProceduresRules,y=log(TotalIncidents+1))) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="SchoolProceduresRules vs. y=log(TotalIncidents+1)") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))
s2

s3 <- ggplot(df2, aes(x=SchoolPreventionProgram, y=log(TotalIncidents+1))) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="SchoolPreventionProgram vs. y=log(TotalIncidents+1)") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))
s3

s4 <- ggplot(df2, aes(x=SchoolTrainingProgram, y=log(TotalIncidents+1))) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="SchoolTrainingProgram vs. y=log(TotalIncidents+1)") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))
s4

ggarrange(s1, s2, s3, s4, ncol=2, nrow=2)
```

```{r}
g1 <- ggplot(df2, aes(factor(SchoolSecurity), log(TotalIncidents+1), fill=factor(SchoolSecurity))) + 
  geom_boxplot() +
  labs(x = "SchoolSecurity", y = "log(TotalIncidents+1)", title = "Boxplot of log(TotalIncidents+1) vs. SchoolSecurity") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g1

g2 <- ggplot(df2, aes(factor(SchoolProceduresRules), log(TotalIncidents+1), fill=factor(SchoolProceduresRules))) + 
  geom_boxplot() +
  labs(x = "SchoolProceduresRules", y = "log(TotalIncidents+1)", title = "Boxplot of log(TotalIncidents+1) vs. SchoolProceduresRules") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g2

g3 <- ggplot(df2, aes(factor(SchoolPreventionProgram), log(TotalIncidents+1), fill=factor(SchoolPreventionProgram))) + 
  geom_boxplot() +
  labs(x = "SchoolPreventionProgram", y = "log(TotalIncidents+1)", title = "Boxplot of log(TotalIncidents+1) vs. SchoolPreventionProgram") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g3

g4 <- ggplot(df2, aes(factor(SchoolTrainingProgram), log(TotalIncidents+1), fill=factor(SchoolTrainingProgram))) + 
  geom_boxplot() +
  labs(x = "SchoolTrainingProgram", y = "TotalIncidents", title = "Boxplot of log(TotalIncidents+1) vs. SchoolTrainingProgram") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g4

ggarrange(g1, g2, g3, g4, ncol=2, nrow=2)
```

## 4. Data Trasformation
## Convert to dummies to be used in next steps

- Create a new dataframe
```{r}
df3<-df2
```

- Convert categorical variables to dummies
```{r}
for(level in unique(df2$StudentAreaCrime)){df3[paste("RacialTensions", level, sep = "_")] <- ifelse(df3$RacialTensions == level, 1, 0)}

for(level in unique(df2$StudentAreaCrime)){df3[paste("StudentBullying", level, sep = "_")] <- ifelse(df3$StudentBullying == level, 1, 0)}

for(level in unique(df2$StudentAreaCrime)){df3[paste("StudentSexualHarassment", level, sep = "_")] <- ifelse(df3$StudentSexualHarassment == level, 1, 0)}

for(level in unique(df2$StudentAreaCrime)){df3[paste("VerbalAbuse", level, sep = "_")] <- ifelse(df3$VerbalAbuse == level, 1, 0)}

for(level in unique(df2$StudentAreaCrime)){df3[paste("SexualOrientationHarassment", level, sep = "_")] <- ifelse(df3$SexualOrientationHarassment == level, 1, 0)}

for(level in unique(df2$StudentAreaCrime)){df3[paste("WidespreadDisorder", level, sep = "_")] <- ifelse(df3$WidespreadDisorder == level, 1, 0)}

for(level in unique(df2$StudentAreaCrime)){df3[paste("GenderIdentityHarassment", level, sep = "_")] <- ifelse(df3$GenderIdentityHarassment == level, 1, 0)}

for(level in unique(df2$StudentAreaCrime)){df3[paste("ActsofDisrespect", level, sep = "_")] <- ifelse(df3$ActsofDisrespect == level, 1, 0)}

for(level in unique(df2$StudentAreaCrime)){df3[paste("GangActivities", level, sep = "_")] <- ifelse(df3$GangActivities == level, 1, 0)}

for(level in unique(df2$StudentAreaCrime)){df3[paste("StudentAreaCrime", level, sep = "_")] <- ifelse(df3$StudentAreaCrime == level, 1, 0)}

for(level in unique(df2$SchoolAreaCrime)){df3[paste("SchoolAreaCrime", level, sep = "_")] <- ifelse(df3$SchoolAreaCrime == level, 1, 0)}

for(level in unique(df3$Urbanicity)){df3[paste("Urbanicity", level, sep = "_")] <- ifelse(df3$Urbanicity == level, 1, 0)}

for(level in unique(df3$GradeLevel)){df3[paste("GradeLevel", level, sep = "_")] <- ifelse(df3$GradeLevel == level, 1, 0)}

for(level in unique(df3$SchoolSize)){df3[paste("SchoolSize", level, sep = "_")] <- ifelse(df3$SchoolSize == level, 1, 0)}

for(level in unique(df3$PctWhiteStudents)){df3[paste("PctWhiteStudents", level, sep = "_")] <- ifelse(df3$PctWhiteStudents == level, 1, 0)}
```

- Remove categorical columns
```{r}
df3[,5:13]<-NULL # Order column looks like meaningless
df3[,11:12]<-NULL # Order column looks like meaningless
df3[,15:18]<-NULL # Order column looks like meaningless
```

- Remove Arrest columns
```{r}
df3[,3]<-NULL # Order column looks like meaningless
```

```{r}
str(df3)
```

## 4. Correlation analysis
- Build correlation and order by decreasing
```{r}
set.seed(123)
library(dplyr)
library(tidyr)

cor(df3) %>%
  as.data.frame() %>%
  mutate(var1 = rownames(.)) %>%
  gather(var2, value, -var1) %>%
  arrange(desc(value)) %>%
  group_by(value) %>%
  filter(row_number()==1)
```

- Remove GenderIdentityHarassment_1 column as a result of collinearity
```{r}
df3[,45]<-NULL # Order column looks like meaningless
str(df3)
```

- Build correlation atts with dependent variable
```{r}
correlations <- cor(df3)
outcome_cor <- correlations[,"TotalIncidents"]
summary(outcome_cor)
```

- correlation atts - Keep top 10 variables with the highest correlation
```{r}
# correlation atts - price - in order
absoutcome_cor <- abs(outcome_cor)
head(absoutcome_cor[order(absoutcome_cor, decreasing = TRUE)],11)
```

- correlation matrix
```{r}
top10 = c("TotalIncidents", "SchoolSize_4", "GangActivities_4", "GangActivities_3", "GradeLevel_3", "GradeLevel_1", "StudentAreaCrime_3", "SchoolAreaCrime_3", "GenderIdentityHarassment_4", "PctWhiteStudents_4", "VerbalAbuse_1")
dftop = df3[,top10]
cortop <- cor(dftop)
corrplot(cortop, method = "color", addCoef.col = "black",number.cex = .5,
         tl.col = "black", tl.srt = 90, diag = FALSE, tl.cex =0.9)
```

## 5. Modeling
## 5.1. Model 1 with top 10 highest correlation with TotalIncidents
```{r}
lm1 <- lm(log(TotalIncidents+1) ~ SchoolSize_4 + GangActivities_4 + GangActivities_3 + GradeLevel_3 + GradeLevel_1 + StudentAreaCrime_3 + SchoolAreaCrime_3 + GenderIdentityHarassment_4 + PctWhiteStudents_4 + VerbalAbuse_1, data = df3)
summary(lm1)
```

Using K-Fold CV
```{r}
# Define training control
set.seed(123) 
train.control <- trainControl(method = "cv", number = 10) #cv Cross-Validation
# Train the model
model1 <- train(log(TotalIncidents+1) ~ SchoolSize_4 + GangActivities_4 + GangActivities_3 + GradeLevel_3 + GradeLevel_1 + StudentAreaCrime_3 + SchoolAreaCrime_3 + GenderIdentityHarassment_4 + PctWhiteStudents_4 + VerbalAbuse_1, data = df3, method = "lm", trControl = train.control)
# Summarize the results
summary(model1)
print(model1)
```

```{r}
ggqqplot(data=data.frame(residuals=residuals(model1)), 
         x="residuals", color="dodgerblue4", ggtheme = theme_pubclean())

iworst1 <- which.max(abs(residuals(model1)))
paste("predicted Total Incidents: ", round(exp(fitted(model1)[iworst]),3), ".", sep="")
df3[iworst1,]
```

## 5.2. LASSO
```{r}
library(glmnet)
x <- model.matrix(TotalIncidents~., df3)[,-1]
y <- df3$TotalIncidents
mod <- cv.glmnet(as.matrix(x), y, alpha=1)
```

To see the coefficients with the minimum cross-validation error
To see the coefficients with the "largest value of lambda such that error is within 1 standard error of the minimum:
```{r}
as.matrix(coef(mod, mod$lambda.min))
as.matrix(coef(mod, mod$lambda.1se))
```

You can also select any other value of lambda that you want. Coefficients that are 0 have been dropped out of the model
```{r}
CF <- as.matrix(coef(mod, mod$lambda.1se))
CF[CF!=0,]
```


```{r}
#Predict the lambda's value
pred.Lasso = predict(mod, test_matrix, type="response", s="lambda.1se")

#plot of coef
plot(mod$glmnet.fit,xvar ="lambda", label = T)
abline(v=log(c(mod$lambda.min,mod$lambda.1se)),lty=2)
```


Using Model 2 using above independent variables
```{r}
lm2 <- lm(log(TotalIncidents+1) ~ GradeLevel_1 + StudentAreaCrime_3 + SchoolAreaCrime_3 + SchoolSize_4  + VerbalAbuse_1 +  GangActivities_3 + GangActivities_4, data=df3)
summary(lm2)
```

Using K-Fold CV
```{r}
# Define training control
set.seed(123) 
train.control <- trainControl(method = "cv", number = 10) #cv Cross-Validation
# Train the model
model2<- train(log(TotalIncidents+1) ~ GradeLevel_1 + StudentAreaCrime_3 + SchoolAreaCrime_3 + SchoolSize_4 + VerbalAbuse_1 +  GangActivities_3 + GangActivities_4, data = df3, method = "lm",
               trControl = train.control)
# Summarize the results
summary(model2)
print(model2)
```

```{r}
ggqqplot(data=data.frame(residuals=residuals(model2)), 
         x="residuals", color="dodgerblue4", ggtheme = theme_pubclean())

iworst2 <- which.max(abs(residuals(model2)))
paste("predicted Total Incidents: ", round(exp(fitted(model2)[iworst]),3), ".", sep="")
df3[iworst2,]
```

## 5.3. Elastic Net - Model 3
```{r}
y <- df3$TotalIncidents
x <- model.matrix(TotalIncidents~., df3)[,-1]
lambdas <- 10^seq(3, -2, by = -.1)

fit <- glmnet(x, y, alpha = 0.5, lambda = lambdas)
summary(fit)
cv_fit <- cv.glmnet(x, y, alpha = 0.5, lambda = lambdas)
plot(cv_fit)
```

```{r}
dim(x)
```

```{r}
cv_fit$lambda.min
```

```{r}
#Show Crucial Parameter
CF <- as.matrix(coef(cv_fit, cv_fit$lambda.1se))
CF[CF!=0,]

#Predict the lambda's value
pred.elastic = predict(cv_fit, test_matrix, type="response", s="lambda.1se")

#plot of coef
plot(cv_fit$glmnet.fit,xvar ="lambda", label = T)
abline(v=log(c(cv_fit$lambda.min,cv_fit$lambda.1se)),lty=2)
```

```{r}
lm3 <- lm(log(TotalIncidents+1) ~ StudentBullying_2   +   VerbalAbuse_2    +  VerbalAbuse_1  +  GangActivities_3 +  GangActivities_4 + StudentAreaCrime_3  + SchoolAreaCrime_3 + GradeLevel_1    +   SchoolSize_4 + PctWhiteStudents_4, data=df3 )
 summary(lm4)
```

Using K-Fold CV
```{r}
# Define training control
set.seed(123) 
train.control <- trainControl(method = "cv", number = 10) #cv Cross-Validation
# Train the model
model3 <- train(log(TotalIncidents+1) ~ StudentBullying_2   +   VerbalAbuse_2    +  VerbalAbuse_1  +  GangActivities_3 +  GangActivities_4 + StudentAreaCrime_3  + SchoolAreaCrime_3 + GradeLevel_1    +   SchoolSize_4 + PctWhiteStudents_4, data = df3, method = "lm",
               trControl = train.control)
# Summarize the results
summary(model3)
print(model3)
```

```{r}
ggqqplot(data=data.frame(residuals=residuals(model3)), 
         x="residuals", color="dodgerblue4", ggtheme = theme_pubclean())

iworst3 <- which.max(abs(residuals(model3)))
paste("predicted Total Incidents: ", round(exp(fitted(model3)[iworst]),3), ".", sep="")
df3[iworst3,]
```

## 5.4 Random Forest Model 4

```{r}
library(randomForest)
library(mlbench)
library(caret) # use createDataPartition() function 
set.seed(95014)

# partition 
#Create Evaluation Sets
set.seed(123)
n = nrow(df3)
trainIndex = sample(1:n, size = round(0.7*n), replace=FALSE)
```

```{r}
#Creates training and test set from observations 
training = df3[trainIndex,]
testing = df3[-trainIndex,]

model4 <- randomForest(log(TotalIncidents+1) ~ ., training, mtry = 3, 
                         importance = TRUE, na.action = na.omit)
print(model4)
summary(model4)

#Plot the error vs the number of trees graph 
plot(model4) 
```

```{r}
varImp(model4)
varImpPlot(model4,type=2,main = "Model 4", n.var=min(10, nrow(model3$importance)))
```

```{r}
# obtain MSE as of last element in fit$mse
# which should match the output from printout
model4$mse[length(model4$mse)]
# take square root to calculate RMSE for the model
sqrt(model4$mse[length(model3$mse)])
```

```{r}
# now illustrate how to calculate RMSE on test data vs. training data
predValues <- predict(model4,testing)

# we can calculate it directly 
library("ie2misc")
sqrt(mean((log(testing$TotalIncidents+1) -predValues)^2)) #RMSE
mean(abs(log(testing$TotalIncidents+1) -predValues)) #MAE
```

## 6. Evaluation

Evaluation dataframe
```{r}
Model <- c("Model 1", "Model 2", "Model 3", "Model 4")
R_squared <- c(0.4201, 0.4105, 0.4262, 0.5057)
RMSE <- c(1.1422, 1.1519, 1.1364, 1.0160)
MAE <- c(0.8903, 0.9021, 0.8927, 0.7879)
ml <- data.frame(Model, R_squared, RMSE, MAE)
```

Plot
```{r}
library(gridExtra)
p1 <- ggplot(ml, aes(Model, RMSE)) + geom_point(aes(colour = factor(Model), size = 4)) + labs(title="RMSE") + theme(plot.title=element_text(hjust=0.5), axis.title.y = element_blank(),axis.title.x = element_blank(), legend.position="none")
p2 <- ggplot(ml, aes(Model, R_squared)) + geom_point(aes(colour = factor(Model), size = 4)) + labs(title="R-Squared") + theme(plot.title=element_text(hjust=0.5), axis.title.y = element_blank(),axis.title.x = element_blank(), legend.position="none")
p3 <- ggplot(ml, aes(Model, MAE)) + geom_point(aes(colour = factor(Model), size = 4)) + labs(title="MAE") + theme(plot.title=element_text(hjust=0.5), axis.title.y = element_blank(),axis.title.x = element_blank(), legend.position="none")
grid.arrange(p2,p1,p3, ncol=3)
```
